---
title: "Historical horse population in Canada"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = here::here())
library(tidyverse)
```

## Aim

This project explores the historical population of horses in Canada between 1906 and 1972 for each Province.


## Data

```{r load data, include = FALSE}
horse_pop <- read_csv("data/00030067-eng.csv")
horse_pop <- filter(horse_pop,
                    DATE == "At June 1 (x 1,000)",
                    GEO != "Canada")
horse_pop$Value <- horse_pop$Value * 1000
```

Horse population data were sourced from the [Government of Canada's Open Data website](http://open.canada.ca/en/open-data). Specifically, these two sources were used:

- [Horses, number on farms at June 1 and at December 1](http://open.canada.ca/data/en/dataset/43b3a9b3-3842-45e7-8bc8-c4c27b9462ab)
- [Horses, number on farms at June 1, farm value per head and total farm value](http://open.canada.ca/data/en/dataset/b374f60b-9580-44dc-83f6-c0a850c15f30)



## Results

```{r plot horses, echo=FALSE, fig.width=10, fig.height=3}
(horse_plot <- ggplot(horse_pop, aes(x = Ref_Date, y = Value)) +
  geom_point() +
  geom_line() +
  xlab("Year") +
  ylab("Number of horses") +
  ggtitle("Historical number of horses per province in Canada") +
  facet_grid(~GEO) +
  scale_y_continuous(labels = scales::comma)) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

We can see from the visualisation above that Ontario, Saskatchewan and Alberta have had the highest horse populations in Canada. All provinces have had a decline in horse
populations since 1940. This is likely due to the rebound of the Canadian automotive
industry after the Great Depression and the Second World War. An interesting follow-up visualization would be car sales per year for each Province over the time period visualised above to further support this hypothesis. 
